Fastest Paths Between Points in $\R^2$ for a Particle Moving with Constant Speed and Bounded Centripetal Acceleration

### 2013-10-11 particle_definitions.tex
Finish path theorems.

\begin{definition}
  Given a particle, a fastest path $\hat{\gamma}(t)$ between two points, $\bvec{X_1}$ and $\bvec{X_2}$, is a valid path such that $T_f(\hat{\gamma}) \leq T_f(\gamma)$ for all valid paths, $\gamma(t)$, between $\bvec{X_1}$ and $\bvec{X_2}$.
\end{definition}

>>>>

\begin{definition}
  Let $p$ be a particle with initial position $X_0$, a fastest path $\hat{\gamma}(t)$ to $\bvec{X_f}$, is a valid path such that $T_f(\hat{\gamma}) \leq T_f(\gamma)$ for all valid paths $\gamma(t)$ to $\bvec{X_f}$.
\end{definition}

### 2013-10-11 path_theorems.tex
Finish path theorems.

The simplest problem to tackle first is the fastest path from $\bvec{X_1}$ to $\bvec{X_2}$ when centripetal acceleration is unbounded, which means that a particle can change its direction of motion instantaneously. The following theorem makes this notion more precise.

\begin{theorem}
  Let a particle $p$ have bounded tangential acceleration and $\theta(T_0) = \theta_0$. For all $\theta_f \in [0, \pi]$ and $\epsilon > 0$, there exists a valid path $\gamma(t)$, such that $\abs{\theta(T_0(\gamma) + \epsilon)} \to \theta_f$ as $\epsilon \to 0$.
\end{theorem}

@@

\begin{theorem}\label{thm:theta-zero}
  Given a particle $p$, a path $\gamma_1$ from $\bvec{X_1}$ to $\bvec{X_2}$ is a fastest path if, for every other valid path $\gamma_2$ from $\bvec{X_1}$ to $\bvec{X_2}$, $\abs{\theta_1(t)} \le \abs{\theta_2(t)}$ and $v_1(t) \le v_2(t)$.

@@

  Thus, if $\abs{\theta_1(t)} \le \abs{\theta_2(t)}$ and $v_1(t) \le v_2(t)$, then $r_1(t) \le r_2(t)$, and $T_f(\gamma_1) \le T_f(\gamma_2)$, so $\gamma_1$ is a fastest path.

@@

\begin{theorem}\label{thm:line-proof}
  Given a particle $p$ with initial position $\bvec{X_1}$, initial velocity $\bvec{v}(t)$, the fastest path $\hat{\gamma}(t)$ from $\bvec{X_1}$ to $\bvec{X_2}$ is the line segment from $\bvec{X_1}$ to $\bvec{X_2}$.
\end{theorem}

\begin{proof}
  Because of Theorem \ref{thm:theta-zero}, there is a path, $\gamma_1(t)$, such that $\theta_1(t) = 0$ for $t \in (T_0(\gamma_1), T_f(\gamma_1)]$ \ref{thm:theta-zero}

\end{proof}

@@

\begin{corollary}
  Given points $\bvec{X_1}, \bvec{X_f} \in \mathrm{R}^2$ and a particle $p$ whose initial position is $\bvec{X_1}$ which moves under conditions set forth in Theorem \ref{thm:line-proof}, the fastest path $\hat{\gamma}(t)$ which $p$ can trace to $\bvec{X_f}$ is unique.
\end{corollary}
\proof We have already shown in Theorem \ref{thm:line-proof} that any fastest path between $\bvec{X_1}$ and $\bvec{X_f}$ follows the straight line between them. Moreover, we showed that when travelling along the fastest path, the particle must have tangential acceleration of $\bar{a}_t$. Since we have starting position $\vec{X_1}$ and initial speed of $0$, the acceleration of the particle uniquely defines a path for the particle.

There is only a single function of the acceleration $a(t) = \bar{a}_t$ which can be satisfied when the particle is moving along a fastest path, therefore, there is only a single possible fastest path.
\qed

We can also investigate how the particle moves along a fastest path. In particular, we can make some statements about the change in $\theta$, the angle between the particle's velocity vector and the straight line between the particle and it's destination.

\begin{theorem}
A 2-dimensional particle, $p$, traveling along a fastest path, $\hat{\gamma(t)}$, from $X_1$ to $X_2$ has $\frac{d\abs{\theta(t)}}{dt} \leq 0$ for all $t \in [0, T_{f, \gamma}]$, unless $\norm{X_1 - X_2} < 2 R_{min}$.
\end{theorem}

\begin{proof}
Since $\theta(T_{f,\gamma}) = 0$, then if $\frac{d\abs{\theta(t)}}{dt} > 0$ for $t \in [t_1, t_2]$, since $\theta{T_{f,\gamma}} < \theta{t_1}$, then by Lemma ???? $\theta(t_1) = \theta(t_2)$ for some $t_2 \in [t_1, T_{f, \gamma}$. Since all of the time derivatives in (\ref{eq:r-diff}) - (\ref{eq:theta-diff}) are all monotonic for $r \in (0, \infty)$, there is no local maximum for them, so the fastest strategy should be the same, regardless of $r$. In other words, there can not be multiple cases for fastest strategies based on $r$.
\end{proof}

>>>>

The simplest problem to tackle first is the fastest path from $\bvec{X_0}$ to $\bvec{X_f}$ when centripetal acceleration is unbounded, which means that a particle can change its direction of motion instantaneously. The following theorem makes this notion more precise.

\begin{theorem}\label{thm:theta-choose}
  Let $p$ be a particle with bounded tangential acceleration and $\theta(T_0) = \theta_0$. For all $\theta_f \in [0, \pi]$ and $\epsilon > 0$, there exists a valid path $\gamma(t)$, such that $\abs{\theta(T_0(\gamma) + \epsilon)} \to \theta_f$ as $\epsilon \to 0$.
\end{theorem}

@@

\begin{theorem}\label{thm:theta-min}
  Let $p$ be a particle with bounded tangential acceleration and initial position $X_0$. A path $\gamma_1$ to $\bvec{X_f}$ is a fastest path if, for every other valid path $\gamma_2$ to $\bvec{X_f}$, $\abs{\theta_1(t)} \le \abs{\theta_2(t)}$ and $v_1(t) \le v_2(t)$. The fastest path is unique if all the inequalities are strict.

@@

  Thus, if $\abs{\theta_1(t)} \le \abs{\theta_2(t)}$ and $v_1(t) \le v_2(t)$, then $r_1(t) \le r_2(t)$, and $T_f(\gamma_1) \le T_f(\gamma_2)$, so $\gamma_1$ is a fastest path. If all the inequalities are strict, then $T_f(\gamma_1) < T_f(\gamma_2)$, so every other path is longer, and thus $\gamma_1$ is the unique fastest path.

@@

\begin{theorem}\label{thm:line-proof}
  Let $p$ be a particle with bounded tangential acceleration, initial position $\bvec{X_0}$, and initial velocity $\bvec{v}(t)$. The line segment from $\bvec{X_0}$ to $\bvec{X_f}$ with $a_c(t) = \bar{a_c}$ for all $t \in [T_0(\gamma_1), T_f(\gamma_1)]$ is a fastest path $\hat{\gamma}(t)$ to $\bvec{X_f}$. Furthermore, the fastest path for this system is unique.
\end{theorem}

\begin{proof}
  Because of Theorem \ref{thm:theta-choose}, there is a valid path $\gamma_1(t)$, which has $\theta_1(t) = 0$ for $t \in (T_0(\gamma_1), T_f(\gamma_1)]$.

  If we let $a_{t, 1}(t) = \bar{a_t}$ for all $t \in [T_0(\gamma_1), T_f(\gamma_1)]$, then $v_1(t) = \bar{a_t} t$, so for every other valid path $\gamma_2$

  \begin{align*}
    v_2(t)& = \int_{t'=T_0(\gamma_2(t))}^{t} a_{t, 2} dt\\
    & \le \bar{a_t} t\\
    &  \le v_1(t)
  \end{align*}

  Thus, all the conditions are satisfied for Theorem \ref{thm:theta-min}, so $\gamma_1(t)$ is a fastest path. The $\theta$ inequality is strict, so the fastest path is unique. Geometrically, since $\theta(t) = 0$, the path is the line segment from $\bvec{X_0}$ to $\bvec{X_f}$.
\end{proof}

@@

% \begin{corollary}
%   The fastest path given by  is unique.
% \end{corollary}
% \begin{proof}
% Since we have starting position $\vec{X_0}$ and initial speed of $0$, the acceleration of the particle uniquely defines a path for the particle.

% There is only a single function of the acceleration $a(t) = \bar{a}_t$ which can be satisfied when the particle is moving along a fastest path, therefore, there is only a single possible fastest path.
% \end{proof}

% We can also investigate how the particle moves along a fastest path. In particular, we can make some statements about the change in $\theta$, the angle between the particle's velocity vector and the straight line between the particle and it's destination.
